European or American? A re-examination of the transatlantic influence over the IMF

### 2025-09-02 manuscript/manuscript.qmd
Appendix: change section title
Fiscal Conditionality Analysis (#56)
* Add fiscal data

* Add cleaned data

* Add fiscal.R

* adapt f_test helper for fiscal

* save reg models

* save reg table

* T+1 in maths mode

* Add fiscal table to appendix

* Add line to reload data

* Add 1st para to robustness checks

* Fix appendix table counter

* Add 2nd para

* 2nd para rephrased

* Appendix: add 1st para

* Appendix: 2nd para

* fiscal: add haven package for read_dta

* appendix: my suggestion for para 3 and 4

* Appendix: revise para 3 and 4

* Appendix: change section title

* Add footnote to ex

# Fiscal Conditions {#sec-fiscal .appendix}

This appendix extends our analysis of IMF conditionality by examining alternative measures that focus specifically on fiscal policy requirements. Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP. We report RFA at two time horizons: one year after programme approval ($T+1$) and at the programme’s end year. Together, these measures provide a more nuanced understanding of the nature and stringency of IMF conditionality, particularly in the fiscal domain.

@tbl-results-fiscal reveals patterns that largely reinforce our main findings while adding important nuance. Consistent with our baseline results, US influence is significantly associated with fewer overall IMF conditions. This pattern extends to fiscal conditions, where US influence is associated with a substantial reduction, though this effect is only marginally significant ($p < 0.1$), likely due to the much smaller sample size ($N=31$ vs. $N=93$ for total conditions). Interestingly, European influence shows the opposite pattern: it is associated with a significant increase in fiscal conditions ($p < 0.05$), suggesting that European powers may advocate for more extensive fiscal reforms when they have influence over IMF lending decisions. This finding supports arguments in the current literature that Europe, if anything, has more fully internalized neoliberal economic orthodoxy and the principles of the "Washington Consensus" than the United States, and may view austerity-oriented policies as intrinsically beneficial [@lutz2014; @dyson2010].

>>>>

# Fiscal Conditionality Analysis {#sec-fiscal .appendix}

This appendix extends our analysis of IMF conditionality by examining alternative measures that focus specifically on fiscal policy requirements. Following @guimaraes2021, we consider not only the total number of conditions (as in our main analysis), but also the number of *fiscal* conditions and the required fiscal adjustment (RFA)[^13]—the extent of fiscal tightening mandated by IMF programmes, expressed as a percentage of GDP. We report RFA at two time horizons: one year after programme approval ($T+1$) and at the programme’s end year. Together, these measures provide a more nuanced understanding of the nature and stringency of IMF conditionality, particularly in the fiscal domain.

[^13]: To merge these variables with our dataset, aggregation is necessary: our data are at the five-year country level, whereas theirs are at the programme level. We sum the number of fiscal conditions and average the RFA variables.

@tbl-results-fiscal reveals patterns that largely reinforce our main findings while adding important nuance. Consistent with our baseline results, US influence is significantly associated with fewer overall IMF conditions. This pattern extends to fiscal conditions, where US influence is associated with a substantial reduction, though this effect is only marginally significant ($p < 0.1$), likely due to the much smaller sample size ($N=31$ vs. $N=93$ for total conditions). Interestingly, European influence shows the opposite pattern: it is associated with a significant increase in fiscal conditions ($p < 0.05$), suggesting that European powers may advocate for more extensive fiscal reforms when they have influence over IMF lending decisions. This finding supports arguments in the current literature that Europe, if anything, has more fully internalised neoliberal economic orthodoxy and the principles of the "Washington Consensus" than the United States, and may view austerity-oriented policies as intrinsically beneficial [@lutz2014; @dyson2010].
